---
title: "holdem_shuffles"
author: "Mike Alexeev"
date: "2023-07-09"
output: pdf_document
---

Adjust graphing functions from:
Merz, David W., and Peter B. Chi. "Knowing When to Fold'em: A Monte Carlo Exploration of Card Shuffling and How Poker Players Can Gain an Advantage." UNLV Gaming Research & Review Journal 26.1 (2022): 7.

Purpose: to plot results that include the Alexeev, Chi wash shuffle model.

```{r}
library(tictoc)
```

```{r}
riffle_shuffle = function(deck) {
  cut_point = rbinom(n = 1,
                     size = length(deck),
                     prob = 1 / 2)
  left = deck[1:cut_point]
  right = deck[(cut_point + 1):length(deck)]
  riffled_deck = c()
  for (i in 1:length(deck)) {
    A = length(left)
    B = length(right)
    random = runif(1)
    odds_left_heap_drop = A / (A + B)
    if (random < odds_left_heap_drop) {
      card_added = left[A]
      left = left[-A]
      A = A - 1
    } else {
      card_added = right[B]
      right = right[-B]
      B = B - 1
    }
    riffled_deck = c(card_added, riffled_deck)
  }
  return(riffled_deck)
}

many_riffle_shuffles = function(deck, iter) {
  for (i in 1:iter) {
    deck = riffle_shuffle(deck)
  }
  return(deck)
}
```

```{r}
strip_shuffle = function(deck) {
  A = rbinom(n = 1,
             size = length(deck),
             prob = 1 / 4)
  B = rbinom(n = 1,
             size = (length(deck) - A),
             prob = 1 / 3)
  C = rbinom(n = 1,
             size = (length(deck) - A - B),
             prob = 1 / 2)
  D = length(deck) - C - B - A
  strip = c(tail(deck, D), deck[(A + B + 1):(A + B + C)], deck[(A + 1):(A +
                                                                          B)], deck[1:A])
  return(strip)
}
```

```{r}
deck_cut = function(deck) {
  cut_pos = rbinom(n = 1,
                   size = length(deck),
                   prob = 1 / 2)
  cut_deck = c(deck[(cut_pos + 1):length(deck)], deck[1:cut_pos])
  return(cut_deck)
}
```

```{r}
holdem_shuffle = function(deck) {
  shuffled_deck = many_riffle_shuffles(1:52, 2)
  shuffled_deck = strip_shuffle(shuffled_deck)
  shuffled_deck = many_riffle_shuffles(shuffled_deck, 1)
  shuffled_deck = deck_cut(shuffled_deck)
  return(shuffled_deck)
}
```

```{r}
alt_shuffle5 = function(deck){
  shuffled_deck = many_riffle_shuffles(deck,5)
  shuffled_deck = strip_shuffle(shuffled_deck)
  shuffled_deck = deck_cut(shuffled_deck)
  return(shuffled_deck)
}
```

returns a deck cut at a certain probability
```{r}
deck_cut_at_prob = function(deck, cut_odds) {
  z = cut_odds
  cards_cut = rbinom(n = 1,
                     size = length(deck),
                     prob = z)
  
  if (cards_cut == length(deck)) {
    return(deck)
  }
  deck_1 = c()
  for (x in 1:length(deck)) {
    if (x <= cards_cut) {
      deck_1 = append(deck_1, deck[x])
    }
  }
  return(deck_1)
}
```

returns the rest of the deck
```{r}
rest_of_deck = function(deck, deck1) {
  deck_2 = c()
  if (length(deck1) == length(deck)) {
    return(deck_2)
  }
  for (x in (length(deck1) + 1):length(deck)) {
    deck_2 = append(deck_2, deck[x])
  }
  return(deck_2)
}
```

takes about 2 seconds per shuffle move for an experienced dealer
6 riffles and a cut takes 14 seconds
```{r}
better_shuffle = function(deck) {
  shuffled_deck = many_riffle_shuffles(1:52, 7)
  return(shuffled_deck)
}
```

Wash shuffle plus casino shuffle parent function.
This is for plotting, not for the Alexeev, Chi model.
```{r}
wash_and_casino = function(deck) {
  deck = wash(deck, 4)
  deck = casino_shuffle(deck)
  deck = array(deck)
  return(deck)
}
```

odds of the neighbor of a card being that close to the card after the shuffle 
i.e. any card in the decks neighbor has about a 12.45 % chance of being it's neighbor again. 
```{r}
card_neighbors_counts = function(sims, shuffle_type, cards) {
  clumps_odds = list()
  bef =  matrix(0, nrow = 52, ncol = 52)
  before1 = list()
  for (i in 1:cards) {
    before1 = append(before1, list(bef))
  }
  for (i in 1:sims) {
    if (sims / 10000 == 0) {print(sims)}
    shuf = shuffle_type(1:52)
    
    for (x in 1:52) {
      for (z in 1:cards) {
        pos = match(x, shuf)
        if ((pos - z) > 0) {
          before = shuf[pos - z]
          before1[[z]][x, before] = before1[[z]][x, before] + 1
        }
      }
    }
  }
  list_of_maxs = list()
  for (i in 1:cards) {
    vals = c()
    for (x in 2:52) {
      before_val = before1[[i]][x, (x - 1)]
      vals = append(vals, before_val)
    }
    list_of_maxs = append(list_of_maxs, list(vals))
  }
  #names(list_of_maxs) = c(1:cards)
  return(list_of_maxs)
}
```

alt shuffles
```{r}
alt3_shuffle1 = function(deck){
  shuffled_deck = many_riffle_shuffles(deck,2)
  shuffled_deck = strip_shuffle(shuffled_deck)
  shuffled_deck = many_riffle_shuffles(shuffled_deck,1)
  shuffled_deck = deck_cut(shuffled_deck)
  return(shuffled_deck)
}

alt3_shuffle2 = function(deck){
  shuffled_deck = many_riffle_shuffles(deck,1)
  shuffled_deck = strip_shuffle(shuffled_deck)
  shuffled_deck = many_riffle_shuffles(shuffled_deck,2)
  shuffled_deck = deck_cut(shuffled_deck)
  return(shuffled_deck)
}

alt3_shuffle3 = function(deck){
  shuffled_deck = many_riffle_shuffles(deck,3)
  shuffled_deck = strip_shuffle(shuffled_deck)
  shuffled_deck = deck_cut(shuffled_deck)
  return(shuffled_deck)
}

alt3_shuffle4 = function(deck){
  shuffled_deck = strip_shuffle(deck)
  shuffled_deck = many_riffle_shuffles(shuffled_deck,3)
  shuffled_deck = deck_cut(shuffled_deck)
  return(shuffled_deck)
}

alt3_shuffle3 = function(deck){
  shuffled_deck = many_riffle_shuffles(deck,3)
  shuffled_deck = strip_shuffle(shuffled_deck)
  shuffled_deck = deck_cut(shuffled_deck)
  return(shuffled_deck)
}

alt4_shuffle5 = function(deck){
  shuffled_deck = many_riffle_shuffles(deck,4)
  shuffled_deck = strip_shuffle(shuffled_deck)
  
  shuffled_deck = deck_cut(shuffled_deck)
  return(shuffled_deck)
}

alt_shuffle5 = function(deck){
  shuffled_deck = many_riffle_shuffles(deck,5)
  shuffled_deck = strip_shuffle(shuffled_deck)
  shuffled_deck = deck_cut(shuffled_deck)
  return(shuffled_deck)
}
```


shell functions for shuffle processes A and B
```{r}
shuffleA_shell = function(deck){
  shuffled_deck = shuffleA(deck, iterations=1248)
  shuffled_deck = casino_shuffle(shuffled_deck)
  shuffled_deck = array(shuffled_deck)
  return(shuffled_deck)
}

shuffleB_shell = function(deck){
  shuffled_deck = shuffleB(deck, p=1/3, t=4)
  shuffled_deck = casino_shuffle(shuffled_deck)
  shuffled_deck = array(shuffled_deck)
  return(shuffled_deck)
}

shuffleC_shell = function(deck){
  shuffled_deck = shuffleC(deck, iterations=1248)
  shuffled_deck = casino_shuffle(shuffled_deck)
  shuffled_deck = array(shuffled_deck)
  return(shuffled_deck)
}

```


odds of the neighbor of a card being that close to the card after the shuffle
i.e. any card in the decks neighbor has about a 12.45 % chance of being it's neighbor again. 
```{r}
average_neighbor = function(sims, shuffle, cards) {
  maxs = card_neighbors_counts(sims, shuffle, cards)
  before_after_vals = c()
  for (i in 1:cards) {
    val = (sum(unlist(maxs[i])) * 100) / (sims * 51)
    before_after_vals = append(before_after_vals, val)
  }
  names(before_after_vals) = c(1:cards)
  return(before_after_vals)
}

iterations = 10000
cards = 52

tic()
hold = average_neighbor(iterations, holdem_shuffle, cards)
wash_neighbors = average_neighbor(iterations, wash_and_casino, cards)
riffle = average_neighbor(iterations, better_shuffle, cards)
processA = average_neighbor(iterations, shuffleA_shell, cards)
processB = average_neighbor(iterations, shuffleB_shell, cards)
processC = average_neighbor(iterations, shuffleC_shell, cards)
alt3_1 = average_neighbor(iterations,alt3_shuffle1,cards)
alt3_2 = average_neighbor(iterations,alt3_shuffle2,cards)
alt3_3 = average_neighbor(iterations,alt3_shuffle3,cards)
alt3_4 = average_neighbor(iterations,alt3_shuffle4,cards)
alt5 = average_neighbor(iterations, alt_shuffle5, cards)
alt3_3 = average_neighbor(iterations,alt3_shuffle3,cards)
alt4_5 = average_neighbor(iterations,alt4_shuffle5,cards)
alt5 = average_neighbor(iterations,alt_shuffle5,cards)
toc()

wash_neighbors
```

```{r}
data_matrix = c()
for (i in 1:52) {
  data_matrix = append(data_matrix, unlist(wash_neighbors_50_50[i]))
  data_matrix = append(data_matrix, unlist(wash_neighbors_60_40[i]))
  data_matrix = append(data_matrix, unlist(wash_neighbors_70_30[i]))
  data_matrix = append(data_matrix, unlist(wash_neighbors_80_20[i]))
}

data <- data.frame(
  values = data_matrix,
  group = rep(c(1:52), each = 4),
  subgroup = rep(
    c(
      "50/50",
      "60/40",
      "70/30",
      "80/20"
    ),
    times = 52
  )
)

ggplot(
  data,
  aes(
    x = group,
    y = values,
    fill = factor(
      subgroup,
      levels = c(
      "50/50",
      "60/40",
      "70/30",
      "80/20"
      )
    )
    
       )) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Percentages of neighbor cards ending n cards apart after shuffling",
       x = "Cards apart",
       y = "Percentage",
       fill = "shuffle") +
  geom_abline(
    intercept = (100 / 51),
    slope = (-100 / (52 * 51)),
    color = "black",
    linetype = "dashed"
  ) +
  theme(
    plot.title = element_text(size = 20, hjust = .5),
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 15),
    text = element_text(size = 20)
  ) +
  scale_fill_manual(values = c("#F8766D", "#6BAED6", "#2CA25F", "#8856A7"))
```
