---
title: "wash_shuffle"
author: "Mike Alexeev"
date: "2023-03-20"
output: pdf_document
---

```{r}
library(ggplot2)
```


Initialize deck structure
```{r}
init_deck = function(deck) {
  # Random partitions for columns to for left/right vectors
  cut1 <- rbinom(n = 1, size = 52, prob = 0.25)
  cut2 <- rbinom(n = 1,
                 size = (52 - cut1),
                 prob = (1 / 3)) + cut1
  cut3 <- rbinom(n = 1,
                 size = (52 - cut2),
                 prob = 0.5) + cut2
  left = deck[c(rev(1:cut1), (cut2 + 1):cut3)]
  right = deck[c(rev((cut3 + 1):52), (cut1 + 1):cut2)]
  
  # Deck will now be a list of left and right vectors
  deck = list(left = left, right = right)
  return(deck)
}
```

Function to circular roll a vector by n steps.
```{r}
roll = function(vec, n) {
  # Circular roll on vector by n steps
  len = length(vec)
  if (len == n) {
    new_vec = vec # If rolling entire vector, then output unchanged
  }
  else {
    new_vec = c(vec[(n + 1):len],vec[1:n])
  }
  return(new_vec)
}
```

Truncated poisson random variable
```{r}
tpois = function(lambda, max = 52) {
  # Returns truncated Poisson random variable with finite max value
  vec = c(rep(0, max))
  for (x in 1:max) {
    vec[x] = ((lambda ^ x) * exp(-lambda)) / factorial(x)
  }
  denom = sum(vec)
  vec = vec / denom
  out = sample(1:max, size = 1, prob = vec)
  return(out)
}
```

Convert deck from vector form to column form.
```{r}
to_col = function(deck) {
  if (length(deck) == 4) {
    print("Deck already in column form.")
    return()
  }
  
  left = deck$left
  right = deck$right
  left_mid = ceiling(length(left) / 2)
  right_mid = ceiling(length(right) / 2)
  
  # Generate column form
  col1 = rev(left[1:left_mid])
  col2 = rev(right[1:right_mid])
  col3 = left[(left_mid + 1):length(left)]
  col4 = right[(right_mid + 1):length(right)]
  
  new_deck = list(col1 = col1, col2 = col2, col3 = col3, col4 = col4)
  
  # Check for missing or extra values
  sanity_check = sum(new_deck$col1, new_deck$col2, 
                     new_deck$col3, new_deck$col4)
  if (sanity_check != 1378) {
    print("Deck integrity lost at to_col.")
    return()
  }
  else {
    return(new_deck)
  }
}
```

Convert deck from column form to vector form.
```{r}
to_vec = function(deck) {
  if (length(deck) == 2) {
    print("Deck already in vector form.")
    return()
  }
  
  left = c(rev(deck$col1), deck$col3)
  right = c(rev(deck$col2), deck$col4)
  new_deck = list(left = left, right = right)
  
  # Check for missing or extra values
  sanity_check = sum(new_deck$left, new_deck$right)
  if (sanity_check != 1378) {
    print("Deck integrity lost at to_vec.")
    return()
  }
  else {
    return(new_deck)
  }
}
```

Function to slice one grouping of cards from one vector to another.
```{r}
slice = function(donor, receiver, rebalance = FALSE) {
  sanity = sum(donor, receiver)
  
  # Define variables
  len_don = length(donor)
  len_rec = length(receiver)
  n = tpois(lambda = 4, max = (len_don - 2)) # Number of cards to donate
  i = ceiling((len_don - n) / 2) # Index at donor before n cards begin
  mid = ceiling(length(receiver) / 2) # Midpoint of receiver where slice inserts
  
  # Move cards
  if (rebalance == FALSE) {
    slice = donor[(i + 1):(i + n)]
    new_receiver = c(receiver[1:mid],
                     slice,
                     receiver[(mid + 1):len_rec])
    new_donor = c(donor[1:i],
                  donor[(i + n + 1):len_don])
  }
  else {
    slice = donor[(i + 1):(i + n)]
    new_receiver = c(receiver, slice)
    new_donor = c(donor[1:i],
                  donor[(i + n + 1):len_don])
  }
  out = list(donor = new_donor, receiver = new_receiver)
  
  # Check for missing or extra values
  sanity_check = sum(out$donor, out$receiver)
  if (sanity != sanity_check) {
    print("Deck integrity lost at slice.")
    return()
  }
  else {
    return(out)
  }
}
```

Function to gather cards back into a single deck.
```{r}
gather = function(deck) {
  # Inputs list of left and right vectors
  vec1 = deck[[1]]
  vec2 = deck[[2]]
  
  # Slice vectors into columns
  col1 = vec1[ceiling(length(vec1) / 2):length(vec1)]
  col2 = vec1[1:(ceiling(length(vec1) / 2) - 1)]
  col3 = vec2[ceiling(length(vec2) / 2):length(vec2)]
  col4 = vec2[1:(ceiling(length(vec2) / 2) - 1)]
  columns = list(col1, col2, col3, col4)
  
  # Need initial column lengths to generate n variable in while loop
  col_len = list(length(col1), length(col2), length(col3), length(col4))
  
  # Create an empty deck
  deck = c()
  counter = 52
  cols = 4
  
  while (counter != 0) {
    i = rbinom(1, (cols - 1), 0.5) + 1 # More likely to sample inner columns
    len = length(columns[[i]])
    n = rbinom(1, (col_len[[i]] - 1), 0.5) + 1
    
    if (len > n) {
      slice = tail(columns[[i]], n)
      deck = c(slice, deck)
      columns[[i]] = columns[[i]][1:(len - n)]
      counter = counter - n
    }
    else if (len == n) {
      slice = columns[[i]]
      deck = c(slice, deck)
      columns[i] = NULL
      counter = counter - n
      cols = cols - 1
    }
    else if (len < n) {
      slice = columns[[i]]
      deck = c(slice, deck)
      columns[i] = NULL
      counter = counter - len
      cols = cols - 1
    }
  }
  return(deck)
}
```

Parent function for wash shuffle
```{r}
wash = function(deck, cycles) {
  deck = init_deck(deck)
  
  min_cards = 5 # Minimum cards single vector may have
  in_out_dist = c(0.3, 0.7) # Inner shuffle occurs 70% of the time
  
  for (i in 1:cycles) {
    # Left hand movement
    n = tpois(3, max = length(deck$left))
    deck$left = roll(deck$left, n)
    
    slice_count1 = 1
    
    for (j in 1:slice_count1) {
      len_left = length(deck$left)
      len_right = length(deck$right)
    
      # Pick which columns to slice
      deck = to_col(deck) # To column form
      if(len_left <= min_cards) {
        # Left vector is too short
        # Rebalance columns 1 and 2
        sliced = slice(deck$col2, deck$col1, rebalance = TRUE)
        deck$col1 = sliced$receiver
        deck$col2 = sliced$donor
        # Rebalance columns 3 and 4
        sliced = slice(deck$col4, deck$col3, rebalance = TRUE)
        deck$col3 = sliced$receiver
        deck$col4 = sliced$donor
      }
      else if (len_right <= min_cards) {
        # Right vector is too short
        # Rebalance columns 1 and 2
        sliced = slice(deck$col1, deck$col2, rebalance = TRUE)
        deck$col1 = sliced$donor
        deck$col2 = sliced$receiver
        # Rebalance columns 3 and 4
        sliced = slice(deck$col3, deck$col4, rebalance = TRUE)
        deck$col3 = sliced$donor
        deck$col4 = sliced$receiver
      }
      else {
        # Neither vector is too short
        inner = sample(c(0, 1), size = 1, prob = in_out_dist)
        if (inner == TRUE) {
          sliced = slice(deck$col3, deck$col2)
          deck$col2 = sliced$receiver
          deck$col3 = sliced$donor
        }
        else if (inner == FALSE) {
          sliced = slice(deck$col1, deck$col2)
          deck$col1 = sliced$donor
          deck$col2 = sliced$receiver
        }
      }
      deck = to_vec(deck) # Back to vector form
    }
    
    # Right hand movement
    n = tpois(3, max = length(deck$right))
    deck$right = roll(deck$right, n)
    
    slice_count2 = 1
    
    for (j in 1:slice_count2) {
      len_left = length(deck$left)
      len_right = length(deck$right)
    
      # Pick which columns to slice
      deck = to_col(deck) # To column form
      if(len_left <= min_cards) {
        # Left vector is too short
        # Rebalance columns 1 and 2
        sliced = slice(deck$col2, deck$col1, rebalance = TRUE)
        deck$col1 = sliced$receiver
        deck$col2 = sliced$donor
        # Rebalance columns 3 and 4
        sliced = slice(deck$col4, deck$col3, rebalance = TRUE)
        deck$col3 = sliced$receiver
        deck$col4 = sliced$donor
      }
      else if (len_right <= min_cards) {
        # Right vector is too short
        # Rebalance columns 1 and 2
        sliced = slice(deck$col1, deck$col2, rebalance = TRUE)
        deck$col1 = sliced$donor
        deck$col2 = sliced$receiver
        # Rebalance columns 3 and 4
        sliced = slice(deck$col3, deck$col4, rebalance = TRUE)
        deck$col3 = sliced$donor
        deck$col4 = sliced$receiver
      }
      else {
        # Neither vector is too short
        inner = sample(c(0, 1), size = 1, prob = in_out_dist)
        if (inner == TRUE) {
          sliced = slice(deck$col2, deck$col3)
          deck$col2 = sliced$donor
          deck$col3 = sliced$receiver
        }
        else if (inner == FALSE) {
          sliced = slice(deck$col4, deck$col3)
          deck$col3 = sliced$receiver
          deck$col4 = sliced$donor
        }
      }
      deck = to_vec(deck)
    }
  }
  
  deck = gather(deck)
  return(deck)
}
```

Functions for riffle shuffle
```{r cache=TRUE}
riffle_shuffle_mine = function(deck) {
  len = length(deck)
  split = rbinom(n = 1, size = len, prob = 0.5)
  A = deck[1:split]
  B = deck[(split + 1):len]
  deck_new = NULL # Empty deck
  
  for (i in 1:len) {
    # Calculate probability of dropping from A
    p = length(A) / (length(A) + length(B))
    
    # Now simulate a single binomial
    # If true, drop from portion A
    if (rbinom(1, 1, prob = p)) {
      deck_new = c(deck_new, A[length(A)])
      A = A[-length(A)]
    }
    else {
      deck_new = c(deck_new, B[length(B)])
      B = B[-length(B)]
    }
  }
  return(deck_new)
}
```

Box shuffle
```{r cache=TRUE}
box = function(deck) {
  A = rbinom(n = 1, size = length(deck), prob = 1 / 4)
  B = rbinom(n = 1, size = (length(deck) - A), prob = 1 / 3)
  C = rbinom(n = 1, size = (length(deck) - A - B), prob = 1 / 2)
  D = length(deck) - C - B - A
  box = c(tail(deck, D), 
          deck[(A + B + 1):(A + B + C)], deck[(A + 1):(A + B)], deck[1:A])
  return(box)
}
```

Cut shuffle
```{r cache=TRUE}
cut = function(deck) {
  split = rbinom(1, 52, 0.5)
  A = deck[1:split]
  B = deck[(split + 1):52]
  new_deck = c(B, A)
  return(new_deck)
}
```

Casino shuffle: riffle, riffle, strip, riffle, cut, deal
```{r cache=TRUE}
casino_shuffle = function(deck) {
  deck = riffle_shuffle_mine(deck)
  deck = riffle_shuffle_mine(deck)
  deck = box(deck)
  deck = riffle_shuffle_mine(deck)
  deck = cut(deck)
  return(deck)
}
```

Monte carlo simulation for wash shuffle followed by casino shuffle
```{r}
reps = 100000

# Vectors to track previous locations of community cards
card14 = c(rep(0, 52))
card15 = c(rep(0, 52))
card16 = c(rep(0, 52))
card18 = c(rep(0, 52))
card20 = c(rep(0, 52))

# Vector to track neighboring card frequency
neighbors = c(rep(0, 52))

for (rep in 1:reps) {
  deck = 1:52
  deck = wash(deck, 4)
  #deck = casino_shuffle(deck)
  
  # Record previous positions of cards
  card14[deck[14]] = card14[deck[14]] + 1
  card15[deck[15]] = card15[deck[15]] + 1
  card16[deck[16]] = card16[deck[16]] + 1
  card18[deck[18]] = card18[deck[18]] + 1
  card20[deck[20]] = card20[deck[20]] + 1
  
  # Record neighboring card positions
  for (k in 1:52) {
    neighbors[k] = neighbors[k] + sum(diff(deck) == k)
  }
}
deck
neighbors = neighbors / reps
card14 = card14 / reps
card15 = card15 / reps
card16 = card16 / reps
card18 = card18 / reps
card20 = card20 / reps
```

We want to know previous positions of community cards
```{r}
jpeg(file = "card14_plot_wash.jpeg",
     width = 480,
     height = 240)
card14_plot = barplot(
  card14,
  names = 1:52,
  main = "Previous location of card 14",
  xlab = "Nth previous index of card 14",
  ylab = "Probability of coming from Nth index"
)
dev.off()

jpeg(file = "card15_plot_wash.jpeg",
     width = 480,
     height = 240)
card15_plot = barplot(
  card15,
  names = 1:52,
  main = "Previous location of card 15",
  xlab = "Nth previous index of card 15",
  ylab = "Probability of coming from Nth index"
)
dev.off()

jpeg(file = "card16_plot_wash.jpeg",
     width = 480,
     height = 240)
card16_plot = barplot(
  card16,
  names = 1:52,
  main = "Previous location of card 16",
  xlab = "Nth previous index of card 16",
  ylab = "Probability of coming from Nth index"
)
dev.off()

jpeg(file = "card18_plot_wash.jpeg",
     width = 480,
     height = 240)
card18_plot = barplot(
  card18,
  names = 1:52,
  main = "Previous location of card 18",
  xlab = "Nth previous index of card 18",
  ylab = "Probability of coming from Nth index"
)
dev.off()

jpeg(file = "card20_plot_wash.jpeg",
     width = 480,
     height = 240)
card20_plot = barplot(
  card20,
  names = 1:52,
  main = "Previous location of card 20",
  xlab = "Nth previous index of card 20",
  ylab = "Probability of coming from Nth index"
)
dev.off()

jpeg(file = "neighboring_cards.jpeg",
     width = 480,
     height = 240)
neighbors_plot = barplot(
  neighbors,
  names = 1:52,
  main = "Neighboring cards frequencies",
  xlab = "Post wash number of cards separating previously neighboring cards",
  ylab = "Frequency per wash shuffle"
)
dev.off()
```

```{r}
# Set up the layout of the multi-panel plot
par(mfrow = c(3, 2))  # 2 rows and 3 columns of plots

barplot(
  card14,
  names = 1:52,
  main = "Previous location of card 14",
  xlab = "Nth previous index of card 14",
  ylab = "Probability of coming from Nth index",
  col = "#6BAED680"
)

barplot(
  card15,
  names = 1:52,
  main = "Previous location of card 15",
  xlab = "Nth previous index of card 15",
  ylab = "Probability of coming from Nth index",
  col = "#6BAED680"
)

barplot(
  card16,
  names = 1:52,
  main = "Previous location of card 16",
  xlab = "Nth previous index of card 16",
  ylab = "Probability of coming from Nth index",
  col = "#6BAED680"
)

barplot(
  card18,
  names = 1:52,
  main = "Previous location of card 18",
  xlab = "Nth previous index of card 18",
  ylab = "Probability of coming from Nth index",
  col = "#6BAED680"
)

barplot(
  card20,
  names = 1:52,
  main = "Previous location of card 20",
  xlab = "Nth previous index of card 20",
  ylab = "Probability of coming from Nth index",
  col = "#6BAED680"
)

par(mfrow = c(2, 1))

barplot(
  neighbors,
  names = 1:52,
  main = "Neighboring cards frequencies",
  xlab = "Post wash number of cards separating previously neighboring cards",
  ylab = "Frequency per wash shuffle",
  col = heat.colors(52, alpha = 0.5)
)

barplot(
  (neighbors + 0.001),
  names = 1:52,
  main = "Neighboring cards frequencies (log scale)",
  xlab = "Post wash number of cards separating previously neighboring cards",
  ylab = "Frequency per wash shuffle",
  col = heat.colors(52, alpha = 0.5),
  log = "y"
)
```

```{r}
barplot(
  (neighbors + 0.001),
  names = 1:52,
  main = "Neighboring cards frequencies (log scale)",
  xlab = "Post wash number of cards separating previously neighboring cards",
  ylab = "Frequency per wash shuffle",
  col = heat.colors(52, alpha = 0.5),
  log = "y"
)
```


Test bench
```{r}
deck = 1:52
deck = wash(deck, 3)
sum(deck)
```

```{r}
reps = 10000
deck = 1:52
for (rep in 1:reps) {
  pre_wash = deck
  deck = wash(pre_wash, 3)
  deck = casino_shuffle(deck)
}
sum(deck)
```

```{r}
reps = 10000
vec = c()
for (rep in 1:reps) {
  tp = tpois(1, 2)
  vec = c(vec, tp)
}
mean(vec)
vec
```








