---
title: "wash_shuffle"
author: "Mike Alexeev"
date: "2023-03-20"
output: pdf_document
---

```{r warning=FALSE}
library(ggplot2)
```

Define washing motion.
The left hand to moves down col 1, up col 3, and repeat.
The right hand moves up col 4, down col 2, and repeat.
Each hand repeat loop simultaneously 4-5 times.
At the end, top card goes on the bottom in some cases.

Function to shift a vector by n steps.
```{r}
shift = function(vec, n) {
  len = length(vec)
  new_vec = c(vec[(len - n):len], vec[1:(len - n - 1)])
  return(new_vec)
}
```

Function to randomly swap two cards to each others' locations.
```{r}
swap_inner = function(vec1, vec2) {
  # Slice vectors into only inner columns
  vec1_inner = vec1[ceiling(length(vec1) / 2):length(vec1)]
  vec2_inner = vec2[ceiling(length(vec2) / 2):length(vec2)]
  
  len1 = length(vec1_inner)
  len2 = length(vec2_inner)
  
  # Probability of the swap coming from vec1 to vec2
  p = len1 / (len1 + len2)
  condition = sample(0:1, 1, prob = c(p, (1 - p)))
  # Choose which vector to swap from
  if (condition == 1) {
    swap_out = vec1_inner
    swap_in = vec2_inner
  }
  else {
    swap_out = vec2_inner
    swap_in = vec1_inner
  }
  
  # Inner swap card index will be binomial distributed
  # Ensure index will never be first or last element
  i = rbinom(1, (length(swap_out) - 1), 0.5) + 1
  
  # Choose group size (1 to 6 cards)
  # Use truncated Poisson for distribution
  n = rbinom(1, 6, 0.5)
  #n = tpois(4)
  
  # Code if-else for if n=0 then break
  # I cannot get this to work for tpois because sometimes n==0
  # The same problem will persist from swap_outer function as well
  if (n == 0) {
    return()
  }
  else {
    # Swap cards
    if ((i + n) > length(swap_out)) {
      # Then card group is [(i-n):i]
      swap_in = c(swap_in[1:(i - n - 1)],
                  swap_out[(i - n):i],
                  swap_in[(i - n):length(swap_in)])
      swap_out = c(swap_out[1:(i - n - 1)],
                   swap_out[(i + 1):length(swap_out)])
    }
    else {
      # Then card group is [i:n]
      swap_in = c(swap_in[1:(i - 1)],
                  swap_out[i:(i + n)],
                  swap_in[i:length(swap_in)])
      swap_out = c(swap_out[1:(i - 1)],
                   swap_out[(i + n + 1):length(swap_out)])
    }
    
    # Join vector halves back together
    if (condition == 1) {
      # Inner swap was left to right
      vec1_new = c(vec1[1:(ceiling(length(vec1) / 2) - 1)], swap_out)
      vec2_new = c(vec2[1:(ceiling(length(vec2) / 2) - 1)], swap_in)
    }
    else {
      # Inner swap was right to left
      vec1_new = c(vec1[1:(ceiling(length(vec1) / 2) - 1)], swap_in)
      vec2_new = c(vec2[1:(ceiling(length(vec2) / 2) - 1)], swap_out)
    }
    
    left = vec1_new
    right = vec2_new
  }
}
```

```{r}
swap_outer = function(vec1, vec2) {
  # Slice vectors into columns
  vec1_outer = vec1[1:(ceiling(length(vec1) / 2) - 1)]
  vec2_inner = vec2[ceiling(length(vec2) / 2):length(vec2)]
  vec1_inner = vec1[ceiling(length(vec1) / 2):length(vec1)]
  vec2_outer = vec2[1:(ceiling(length(vec2) / 2) - 1)]
  
  len1_outer = length(vec1_outer)
  len2_inner = length(vec2_inner)
  len1_inner = length(vec1_inner)
  len2_outer = length(vec2_inner)
  
  # Probability of the swap coming from vec1 to vec2
  p = len1_outer / (len1_outer + len2_outer)
  condition = sample(0:1, 1, prob = c(p, (1 - p)))
  
  # Choose which vector to swap from
  if (condition == 1) {
    out_to_in = sample(0:1, 1)
    if (out_to_in == 1) {
      swap_out = vec1_outer
      swap_in = vec2_inner
    }
    else if (out_to_in == 0) {
      swap_out = vec2_inner
      swap_in = vec1_outer
    }
  }
  else {
    out_to_in = sample(0:1, 1)
    if (out_to_in == 1) {
      swap_out = vec2_outer
      swap_in = vec1_inner
    }
    else if (out_to_in == 0) {
      swap_out = vec1_inner
      swap_in = vec2_outer
    }
  }
  
  # Inner swap card index will be binomial distributed
  # Ensure index will never be first or last element
  i = rbinom(1, (length(swap_out) - 1), 0.5) + 1
  
  # Choose group size (1 to 6 cards)
  # Use truncated Poisson for distribution
  n = rbinom(1, 2, 0.5)
  
  # Code if-else for if n=0 then break
  
  # Swap cards
  if ((i + n) > length(swap_out)) {
    # Then card group is [(i-n):i]
    swap_in = c(swap_in[1:(i - n - 1)],
                swap_out[(i - n):i],
                swap_in[(i - n):length(swap_in)])
    swap_out = c(swap_out[1:(i - n - 1)],
                 swap_out[(i + 1):length(swap_out)])
  }
  else {
    # Then card group is [i:n]
    swap_in = c(swap_in[1:(i - 1)],
                swap_out[i:(i + n)],
                swap_in[i:length(swap_in)])
    swap_out = c(swap_out[1:(i - 1)],
                 swap_out[(i + n + 1):length(swap_out)])
  }
  
  # Join vector halves back together
  if (condition == 1) {
    # Outer swap on left side
    if (out_to_in == 1) {
      left = c(swap_out, vec1_inner)
      right = c(vec2_outer, swap_in)
    }
    else if (out_to_in == 0) {
      left = c(swap_in, vec1_inner)
      right = c(vec2_outer, swap_out)
    }
  }
  else {
    # Outer swap on right side
    if (out_to_in == 1) {
      left = c(vec1_outer, swap_in)
      right = c(swap_out, vec2_inner)
    }
    else if (out_to_in == 0) {
      left = c(vec1_outer, swap_out)
      right = c(swap_in, vec2_inner)
    }
  }
}
```

Function to gather cards back into a single deck.
```{r}
gather = function(vec1, vec2) {
  # Slice vectors into columns
  col1 = vec1[ceiling(length(vec1) / 2):length(vec1)]
  col2 = vec1[1:(ceiling(length(vec1) / 2) - 1)]
  col3 = vec2[ceiling(length(vec2) / 2):length(vec2)]
  col4 = vec2[1:(ceiling(length(vec2) / 2) - 1)]
  columns = list(col1, col2, col3, col4)
  
  # Create an empty deck
  deck = c()
  counter = 52
  cols = 4
  
  while (counter != 0) {
    i = rbinom(1,3,0.5) + 1 # More likely to sample inner columns
    n = rbinom(1, length(i), 0.5)
    len = length(columns[[i]])
    
    if (n == 0) {
      next
    }
    else if (len > 0) {
      if (len > n) {
        slice = tail(columns[[i]], n)
        deck = c(slice, deck)
        columns[[i]] = columns[[i]][1:(len - n)]
        counter = counter - n
      }
      else if (len == n) {
        slice = columns[[i]]
        deck = c(slice, deck)
        columns[i] = NULL
        counter = counter - n
        cols = cols - 1
      }
      else if (len < n) {
        slice = columns[[i]]
        deck = c(slice, deck)
        columns[i] = NULL
        counter = counter - len
        cols = cols - 1
      }
    }
    else if (columns[[i]] == 0) {
      next
    }
  }
  
  return(deck)
}
```

Truncated poisson random variable
```{r}
tpois = function(lambda, max = 52) {
  vec = c(rep(0, max))
  for (x in 1:max) {
    vec[x] = ((lambda ^ x) * exp(-lambda)) / factorial(x)
  }
  denom = sum(vec)
  vec = vec / denom
  out = sample(1:max, size = 1, prob = vec)
  return(out)
}

reps = 100
vec = c()
for (k in 1:reps) {
  vec = c(vec, tpois(100))
}
vec
```

Functions for riffle shuffle
```{r cache=TRUE}
riffle = function(deck) {
  len = length(deck)
  split = rbinom(n = 1, size = len, prob = 0.5)
  A = deck[1:split]
  B = deck[(split + 1):len]
  deck_new = NULL # Empty deck
  
  for (i in 1:len) {
    # Calculate probability of dropping from A
    p = length(A) / (length(A) + length(B))
    
    # Now simulate a single binomial
    # If true, drop from portion A
    if (rbinom(1, 1, prob = p)) {
      deck_new = c(deck_new, A[length(A)])
      A = A[-length(A)]
    }
    else {
      deck_new = c(deck_new, B[length(B)])
      B = B[-length(B)]
    }
  }
  return(deck_new)
}
```

Box shuffle
```{r cache=TRUE}
box = function(deck) {
  A = rbinom(n = 1, size = length(deck), prob = 1 / 4)
  B = rbinom(n = 1, size = (length(deck) - A), prob = 1 / 3)
  C = rbinom(n = 1, size = (length(deck) - A - B), prob = 1 / 2)
  D = length(deck) - C - B - A
  box = c(tail(deck, D), 
          deck[(A + B + 1):(A + B + C)], deck[(A + 1):(A + B)], deck[1:A])
  return(box)
}
```

Cut shuffle
```{r cache=TRUE}
cut = function(deck) {
  split = rbinom(1, 52, 0.5)
  A = deck[1:split]
  B = deck[(split + 1):52]
  new_deck = c(B, A)
  return(new_deck)
}
```

Casino shuffle: riffle, riffle, strip, riffle, cut, deal
```{r cache=TRUE}
casino_shuffle = function(deck) {
  deck = riffle(deck)
  deck = riffle(deck)
  deck = box(deck)
  deck = riffle(deck)
  deck = cut(deck)
  return(deck)
}
```

Monte carlo simulation for wash shuffle followed by casino shuffle
```{r}
reps = 10000

# Vectors to track previous locations of community cards
card14 = c(rep(0, 52))
card15 = c(rep(0, 52))
card16 = c(rep(0, 52))
card18 = c(rep(0, 52))
card20 = c(rep(0, 52))

# Vector to track neighboring card frequency
neighbors = c(rep(0, 52))

for (rep in 1:reps) {
  # Initialize cards
  left = c(1:13, rev(27:39))
  right = c(rev(40:52), 14:26)
  cycles = 4
  
  for (i in 1:cycles) {
    n = tpois(3) # Random length of shift
    left = shift(left, n)
    
    for (j in 1:5) {
      outer_inner = sample(0:1, 1)
      
      if (outer_inner == 1) {
        swap_inner(left, right)
      }
      else {
        swap_outer(left, right)
      }
    }
    
    n = tpois(3) # Random length of shift
    right = shift(right, n)
    
    for (j in 1:5) {
      outer_inner = sample(0:1, 1)
      
      if (outer_inner == 1) {
        swap_inner(left, right)
      }
      else {
        swap_outer(left, right)
      }
    }
  }
  
  deck = gather(left, right)
  deck = casino_shuffle(deck)
  
  # Record previous positions of cards
  card14[deck[14]] = card14[deck[14]] + 1
  card15[deck[15]] = card15[deck[15]] + 1
  card16[deck[16]] = card16[deck[16]] + 1
  card18[deck[18]] = card18[deck[18]] + 1
  card20[deck[20]] = card20[deck[20]] + 1
  
  # Record neighboring card positions
  for (k in 1:52) {
    neighbors[k] = neighbors[k] + sum(diff(deck) == k)
  }
}
deck
```

We want to know previous positions of community cards
```{r}
jpeg(file = "card14_plot_wash_and_cs.jpeg",
     width = 480,
     height = 240)
card14_plot = barplot(card14, names = 1:52, main = "Previous location of card 14")
dev.off()

jpeg(file = "card15_plot_wash_and_cs.jpeg",
     width = 480,
     height = 240)
card15_plot = barplot(card15, names = 1:52, main = "Previous location of card 15")
dev.off()

jpeg(file = "card16_plot_wash_and_cs.jpeg",
     width = 480,
     height = 240)
card16_plot = barplot(card16, names = 1:52, main = "Previous location of card 16")
dev.off()

jpeg(file = "card18_plot_wash_and_cs.jpeg",
     width = 480,
     height = 240)
card18_plot = barplot(card18, names = 1:52, main = "Previous location of card 18")
dev.off()

jpeg(file = "card20_plot_wash_and_cs.jpeg",
     width = 480,
     height = 240)
card20_plot = barplot(card20, names = 1:52, main = "Previous location of card 20")
dev.off()

card14_plot = barplot(card14, names = 1:52, main = "Previous location of card 14")
card15_plot = barplot(card15, names = 1:52, main = "Previous location of card 15")
card16_plot = barplot(card16, names = 1:52, main = "Previous location of card 16")
card18_plot = barplot(card18, names = 1:52, main = "Previous location of card 18")
card20_plot = barplot(card20, names = 1:52, main = "Previous location of card 20")
neighbors_plot = barplot(neighbors, names = 1:52, main = "Neighboring cards frequencies")
```

Monte Carlo simulation for only casino shuffle
```{r}
reps = 100000

# Vectors to track previous locations of community cards
card14_casino = c(rep(0, 52))
card15_casino = c(rep(0, 52))
card16_casino = c(rep(0, 52))
card18_casino = c(rep(0, 52))
card20_casino = c(rep(0, 52))

# Vector to track neighboring card frequency
neighbors_casino = c(rep(0, 52))

for (rep in 1:reps) {
  deck = 1:52
  deck = casino_shuffle(deck)
  
  # Record previous positions of cards
  card14_casino[deck[14]] = card14_casino[deck[14]] + 1
  card15_casino[deck[15]] = card15_casino[deck[15]] + 1
  card16_casino[deck[16]] = card16_casino[deck[16]] + 1
  card18_casino[deck[18]] = card18_casino[deck[18]] + 1
  card20_casino[deck[20]] = card20_casino[deck[20]] + 1
  
  # Record neighboring card positions
  for (k in 1:52) {
    neighbors_casino[k] = neighbors_casino[k] + sum(diff(deck) == k)
  }
}
```

Generate dataframe
```{r}
data = data.frame(
  cards = (1:52),
  neighbors = neighbors,
  neighbors_casino = neighbors_casino
)
```

ggplots
```{r}
ggplot(data, aes(x = cards)) + 
  geom_bar(aes(y = neighbors_casino, fill = "Vector 2"), stat = "identity") +
  scale_fill_manual(name = "Vector",
                    values = c("Vector 1" = "turquoise3", "Vector 2" = "tomato"))
```












Debugger
```{r}
func = swap_inner
debug(func)
left = c(1:13, rev(27:39))
right = c(rev(40:52), 14:26)
swap_inner(left, right)
undebug(func)
```



Metrics:
-take a look at graphs from the paper
-evaluate what happens in a poker hand after a shuffle (with 6 players)
-figure out what cards are in the dealt positions

1. Fix direction of shift function. Done

2. Code v2 of swap function. Done
  Must account for more than one card in swap
  Not necessarily always a 1-for-1 swap
  Left and right vectors not always the same length

3. Code v2 of gather function where: Done
  Rebuild the deck from the bottom to the top of the two inside columns
  two inside columns mostly stay together
  two outside columns more likely to go in deck at random as singles

4. Add riffle, with box and cut. Done

5. Add Poisson truncated for number of cards in swap and in shift. Done

6. Replicate graphs from original paper with wash in front.













