---
title: "white_paper_shuffles"
author: "Mike Alexeev"
output: pdf_document
date: "2024-11-30"
---

```{r}
library(ggplot2)
```

Need to model white paper shuffles
Link to paper:
https://arxiv.org/pdf/1910.02389

Shuffle process A

-Cards begins by occupying spaces 1 through n, where n is the total number of cards.
-At each step, choose card randomly.
-Chosen card moves left one step, right one step, or stays in place, with probabilities 0.25, 0.25, and 0.5 respectively.
-The step size is calculated randomly.
-Chosen card moves to that slot and other cards slide to make room.
-This process is repeated for some specified number of steps.

```{r}
# Function to place a card in a pile at any spot if cards are already there
place_card <- function(spot, new_card){
  if(length(spot) == 1){
    return(c(spot, new_card)[sample(2)])  # return the 2 cards in random order
  } else{
    placement <- sample(length(spot)+1, size=1)
    if(placement == 1){
      return(c(new_card, spot))
    } else if(placement == (length(spot) + 1)) {
      return(c(spot, new_card))
    } else{
      return(c(spot[1:(placement-1)], new_card, spot[placement:length(spot)]))
    }
  }
}

# Function to move a card to a new location
make_move <- function(deck){
  # pick card, not location
  to_move <- sample(52, size=1)
  
  # now get the card's location
  ind <- which(sapply(deck, function(e) is.element(to_move, e)))

  # determine if and where it will move to -- 4/8/25 still in progress here
  move <- sample(c(-1, 0, 1), size=1, prob = c(0.25, 0.5, 0.25))
  
  # Do nothing if it doesn't move, otherwise move it (not done yet)
  if(move !=0){
    if(ind == 1 & move == -1){
      new_ind <- 52
    } else if(ind == 52 & move == 1){
      new_ind <- 1
    } else {
      new_ind <- ind + move
    }

    deck[[new_ind]] <- place_card(deck[[new_ind]], to_move)
    deck[[ind]] <- deck[[ind]][-which(deck[[ind]]==to_move)]
    # need to remove card from where it came, too
  }  
  return(deck)
}

# Wrapper function
# its input should be a deck in vector format
shuffleA <- function(deck, iterations){
  deck <- as.list(deck)
  for(i in 1:iterations){
    deck <- make_move(deck)
  }
  return(unlist(deck))
}

set.seed(12)
deck <- 1:52
shuffleA(deck, iterations=1248)


```

Shuffle process B

-Cards begins by occupying spaces 1 through n, where n is the total number of cards.
-At each step, every card moves a random number of spaces to the right according to independent geometric random variables.
-Then every card moves a random number of spaces to the left according to the same distribution.
-Cards cannot go past positions 1 and 52.
-When cards enter an occupied position, they are merged with the cards already there by a GSR shuffle.
-This is repeated for some specified number of steps, and then all piles are gathered together in order.

```{r}
# its input should be a deck in vector format
# Following Remark 26: move cards one at a time, inserting them at a random position
# in the pile at their destination

shuffleB <- function(deck, p, t){
  n <- length(deck)
  deck <- as.list(deck)
  for(i in 1:t){
    # sweep to the right
    moves <- (rgeom(n, p) + 1) # rgeom in R is failures; paper defines it as trials
    for(j in 1:n){  
      ind <- which(sapply(deck, function(e) is.element(j, e))) # get the location of the card to move
      new_ind <- ind + moves[j] # go to the right
      if(new_ind > 52){
        new_ind <- 52 # do not go past 52
      } 
      
      if(new_ind != ind){
        deck[[new_ind]] <- place_card(deck[[new_ind]], j) # Remark 26: insert at random position
        deck[[ind]] <- deck[[ind]][-which(deck[[ind]]==j)] # remove it from its previous position
      }
    }
    
    # sweep to the left
    moves <- (rgeom(n, p) + 1) # rgeom in R is failures; paper defines it as trials
    for(j in 1:n){  
      ind <- which(sapply(deck, function(e) is.element(j, e)))
      new_ind <- ind - moves[j] # go to the left
      if(new_ind < 1){
        new_ind <- 1 # do not go past 1
      }
      
      if(new_ind != ind){
        deck[[new_ind]] <- place_card(deck[[new_ind]], j)
        deck[[ind]] <- deck[[ind]][-which(deck[[ind]]==j)]
      }
    }  

  }
  return(unlist(deck))
}

set.seed(12)
deck <- 1:52
shuffleB(deck, p=1/3, t=4)
```

```{r}
reps = 1000

# Vectors to track previous locations of community cards
card14 = c(rep(0, 52))
card15 = c(rep(0, 52))
card16 = c(rep(0, 52))
card18 = c(rep(0, 52))
card20 = c(rep(0, 52))

# Vector to track neighboring card frequency
neighbors = c(rep(0, 52))

for (rep in 1:reps) {
  #deck = shuffle_process_a(deck = 1:52, reps = 100)
  #deck = shuffle_process_b(deck = 1:52, n = 52)
  deck = shuffle_process_c(num_cards = 52, dims = 3, reps = 100, grid_size = 5)
  #deck = casino_shuffle(deck = 1:52)
  
  # Record previous positions of cards
  card14[deck[14]] = card14[deck[14]] + 1
  card15[deck[15]] = card15[deck[15]] + 1
  card16[deck[16]] = card16[deck[16]] + 1
  card18[deck[18]] = card18[deck[18]] + 1
  card20[deck[20]] = card20[deck[20]] + 1
  
  # Record neighboring card positions
  for (k in 1:52) {
    neighbors[k] = neighbors[k] + sum(diff(deck) == k)
  }
}
deck
neighbors = neighbors / reps
card14 = card14 / reps
card15 = card15 / reps
card16 = card16 / reps
card18 = card18 / reps
card20 = card20 / reps
```

Shuffle process C

```{r}
# Shuffle Process C
# Specifically for d=2

# Function to place a card in a pile at any spot if cards are already there
place_card <- function(spot, new_card){
  if(length(spot) == 1){
    return(c(spot, new_card)[sample(2)])  # return the 2 cards in random order
  } else{
    placement <- sample(length(spot)+1, size=1)
    if(placement == 1){
      return(c(new_card, spot))
    } else if(placement == (length(spot) + 1)) {
      return(c(spot, new_card))
    } else{
      return(c(spot[1:(placement-1)], new_card, spot[placement:length(spot)]))
    }
  }
}


# Function to move a card to a new location
make_moveC <- function(deck){
  # pick card, not location
  to_move <- sample(52, size=1)
  
  # now get the card's location
  ind1 <- NA
  ind2 <- NA
  i <- 1
  j <- 1
  
  while(is.na(ind1) | is.na(ind2)){
    if(is.element(to_move, deck[[i]][[j]])){
      ind1 <- i
      ind2 <- j
    }
    i <- ifelse(j==52, i+1, i)
    j <- ifelse(j==52, 1, j+1)
  }
  
#  ind1 <- which(sapply(deck, function(e) is.element(to_move, e)))
#  ind2 <- which(sapply(deck[[ind1]], function(e) is.element(to_move, e)))

  ind1x <- ind1
  ind2x <- ind2
  
  # can keep it like this, and then 50/50 chance of moving vertically or horizontally
  # sort of (with adjustments as below)
  move <- sample(c(-1, 0, 1), size=1, prob = c(0.25, 0.5, 0.25))
  
  # Do nothing if it doesn't move, otherwise move it
  if(move !=0){
    # pick which direction it will move in
    # if the card is not on an edge of the grid, this is all we need to do
    # otherwise we need to adjust probabilities
    dir <- sample(c(1,2), size=1)
    if(dir == 1){
      ind1 <- ind1 + move
      if(ind1 == 0){
        if(ind2 == 1 | ind2==52){
          # this means the card was in a corner of the grid, meaning
          # there are only 2 neighbors
          # In this case, just move ind1 the other way
          # This will occur with overall probability 0.5*0.5 as it should
          ind1 <- 2
        } else{
          # otherwise, need to balance probability for two ind2 options
          # if we get to here, this means that the card is on an edge, but not in a corner
          # all moves should have overall probability 1/6 
          # The probability of getting here is 0.5*0.5 = 1/4, so we need to adjust that down to 1/6
          # 1/4 * 2/3 = 1/6
          adj <- sample(c(1,2), size=1, prob=c(2/3, 1/3))
          if(adj==1){
            ind1 <- 2
          } else {
            # with 1/3 probability, we do not move ind1,
            # and send it over to move in the other direction
            dir <- 2
            # and move ind1 back
            ind1 <- ind1 - move
          }
        }
      }
      if(ind1 == 53){
        if(ind2 == 1 | ind2==52){
          # in this case, just move ind1 the other way
          ind1 <- 51
        } else {
          # otherwise, again need to balance probability
          adj <- sample(c(1,2), size=1, prob=c(2/3, 1/3))
          if(adj==1){
            ind1 <- 51
          } else {
            dir <- 2
            # and move ind1 back
            ind1 <- ind1 - move
          }
        }
      }
    }
    if(dir==2){
      ind2 <- ind2 + move
      if(ind2 == 0){
        ind2 <- 2
      }
      if(ind2 == 53){
        ind2 <- 51
      }
    }
    # I think this works -- all moves should have the correct probability

    # need to remove card from where it came
    deck[[ind1]][[ind2]] <- place_card(deck[[ind1]][[ind2]], to_move)
    deck[[ind1x]][[ind2x]] <- deck[[ind1x]][[ind2x]][-which(deck[[ind1x]][[ind2x]]==to_move)]
    
  }  
  return(deck)
}


# Wrapper function
# its input should be a deck in vector format
shuffleC <- function(deck, iterations){
  # 2-dimensional list
  deck <- list(as.list(deck))
  for(i in 2:52){
    deck[[i]] <- as.list(rep(NA, 52))
    # inefficient but I don't see any other way
    for(j in 1:52){
      deck[[i]][[j]] <- integer(0)
    }
  }
  for(i in 1:iterations){
    deck <- make_moveC(deck)
  }
  return(unlist(deck))
}

```
