---
title: "white_paper_shuffles"
author: "Mike Alexeev"
output: pdf_document
date: "2024-11-30"
---

```{r}
library(ggplot2)
```

Need to model white paper shuffles
Link to paper:
https://arxiv.org/pdf/1910.02389

Shuffle process A

-Cards begins by occupying spaces 1 through n, where n is the total number of cards.
-At each step, choose card randomly.
-Chosen card moves left one step, right one step, or stays in place, with probabilities 0.25, 0.25, and 0.5 respectively.
-The step size is calculated randomly.
-Chosen card moves to that slot and other cards slide to make room.
-This process is repeated for some specified number of steps.

```{r}
shuffle_process_a = function(new_deck, reps) {
  # Loop to randomly move individual cards
  for (rep in 1:reps) {
    # Pull in a new deck
    deck = new_deck
    
    # Pick a card and direction of travel
    card = sample(deck, 1)
    dir = sample(c(-1, 0, 1), 1, prob = c(0.25, 0.25, 0.5))
    
    # Initialize new location parameter
    new_loc = 0
    
    # Choose new location for chosen card
    if (dir == -1) {
      new_loc = sample(c(1:card - 1), 1)
    } else if (dir == 1) {
      new_loc = sample(c(card + 1:52), 1)
    }
    
    # Initialize blank deck
    new_deck = NULL
    
    # Generate deck using new card placement
    if (new_loc == 0) {
      new_deck = deck
    } else if (new_loc == 1) {
      new_deck = c(card, deck[!deck == card])
    } else if (new_loc == 52) {
      new_deck = c(deck[!deck == card], card)
    } else {
      new_deck = append(deck[!deck == card], card, after = new_loc - 1)
    }
  }
  return(new_deck)
}
# Check results
new_deck = shuffle_process_a(1:52, 100)
print(sum(new_deck) == sum(1:52))
print(new_deck)
```

Shuffle process B

-Cards begins by occupying spaces 1 through n, where n is the total number of cards.
-At each step, every card moves a random number of spaces to the right according to independent geometric random variables.
-Then every card moves a random number of spaces to the left according to the same distribution.
-When cards enter an occupied position, they are merged with the cards already there by a GSR shuffle.
-This is repeated for some specified number of steps, and then all piles are gathered together in order.

```{r}
shuffle_process_b = function(deck, n = 52) {
  shift = rgeom(n = n, prob = 0.25) - rgeom(n = n, prob = 0.25)
  deck = deck + shift
  
  # Initialize new_deck to store results
  new_deck = numeric(n)
  
  # Track the remaining indices of the deck
  remaining_indices = 1:n
  
  # Loop through each position in the new deck
  for (i in 1:n) {
    # Get the index of the smallest card in the remaining deck
    min_index = remaining_indices[which.min(deck[remaining_indices])]
    
    # Get the corresponding card and adjust it by the shift
    card = deck[min_index] - shift[min_index]
    
    # Store the adjusted card in the new_deck
    new_deck[i] = card
    
    # Remove the processed card from the remaining_indices
    remaining_indices = remaining_indices[remaining_indices != min_index]
  }
  return(new_deck)
}
# Check results
new_deck = shuffle_process_b(1:52, 52)
print(sum(new_deck) == sum(1:52))
print(new_deck)

```


Shuffle process C

-This is a generalization of Shuffling Process A into multiple dimensions.
-For some number of dimensions d, cards begin by occupying arbitrary vertices in a d-dimensional grid.
-At each step, choose a random card.
-With probability ½, it remains at its current vertex.
-Otherwise, with equal probability it will move to any neighboring vertex.
-If a card moves to an occupied position, it is merged with the cards already there at a random position. 
-This is repeated for some specified number of steps, and then all piles are gathered together in lexicographical order.

```{r}
# Initialize the positions of cards randomly across a d-dimensional bounded grid
initialize_positions = function(num_cards, dims, grid_size) {
  # Generate positions for cards randomly within a bounded grid
  positions = matrix(sample(1:grid_size, num_cards * dims, replace = TRUE), ncol = dims)
  return(positions)
}

# Get neighboring positions in a d-dimensional bounded grid
get_neighbors = function(pos, grid_size) {
  neighbors = list()
  dims = length(pos)
  for (i in 1:dims) {
    for (delta in c(-1, 1)) {
      neighbor = pos
      neighbor[i] = neighbor[i] + delta
      # Ensure neighbors are within the grid bounds
      if (all(neighbor >= 1 & neighbor <= grid_size)) {
        neighbors = append(neighbors, list(neighbor))
      }
    }
  }
  return(neighbors)
}

shuffle_process_c = function(num_cards, dims, reps, grid_size) {
  # Generate starting positions randomly
  positions = initialize_positions(num_cards, dims, grid_size)
  
  # Piles vector creates a list for each starting card
  # Used to track the random positioning of overlapping card vertices
  piles = vector("list", grid_size^dims)
  
  # Assign each card to its initial pile to track overlaps
  for (i in 1:num_cards) {
    index = paste(positions[i, ], collapse = ",")
    if (is.null(piles[[index]])) {
      piles[[index]] = list()
    }
    piles[[index]] = append(piles[[index]], i)
  }
  
  # Perform shuffling
  for (rep in 1:reps) {
    card_index = sample(1:num_cards, 1)
    current_position = positions[card_index, ]
    
    if (runif(1) > 0.5) {
      # Card moves to a neighboring position
      neighbors = get_neighbors(current_position, grid_size)
      new_position = sample(neighbors, 1)[[1]]
      
      # Update piles
      current_index = paste(current_position, collapse = ",")
      new_index = paste(new_position, collapse = ",")
      
      # Remove card from current pile
      piles[[current_index]] = setdiff(piles[[current_index]], card_index)
      
      # Add card to new pile, inserting at a random position
      if (is.null(piles[[new_index]])) {
        piles[[new_index]] = list()
      }
      piles[[new_index]] = append(
        piles[[new_index]],
        list(card_index),
        after = sample(0:length(piles[[new_index]]), 1)
      )
      
      # Update card position
      positions[card_index, ] = new_position
    }
  }
  
  # Combine piles in lexicographic order
  all_piles = list()
  for (index in names(piles)) {
    if (!is.null(piles[[index]])) {
      all_piles = c(all_piles, unlist(piles[[index]]))
    }
  }
  
  return(unlist(all_piles))
}

# Check results
new_deck = shuffle_process_c(num_cards = 52, dims = 3, reps = 100, grid_size = 5)
print(sum(new_deck) == sum(1:52))
print(new_deck)

```

```{r}
reps = 1000

# Vectors to track previous locations of community cards
card14 = c(rep(0, 52))
card15 = c(rep(0, 52))
card16 = c(rep(0, 52))
card18 = c(rep(0, 52))
card20 = c(rep(0, 52))

# Vector to track neighboring card frequency
neighbors = c(rep(0, 52))

for (rep in 1:reps) {
  #deck = shuffle_process_a(deck = 1:52, reps = 100)
  #deck = shuffle_process_b(deck = 1:52, n = 52)
  deck = shuffle_process_c(num_cards = 52, dims = 3, reps = 100, grid_size = 5)
  #deck = casino_shuffle(deck = 1:52)
  
  # Record previous positions of cards
  card14[deck[14]] = card14[deck[14]] + 1
  card15[deck[15]] = card15[deck[15]] + 1
  card16[deck[16]] = card16[deck[16]] + 1
  card18[deck[18]] = card18[deck[18]] + 1
  card20[deck[20]] = card20[deck[20]] + 1
  
  # Record neighboring card positions
  for (k in 1:52) {
    neighbors[k] = neighbors[k] + sum(diff(deck) == k)
  }
}
deck
neighbors = neighbors / reps
card14 = card14 / reps
card15 = card15 / reps
card16 = card16 / reps
card18 = card18 / reps
card20 = card20 / reps
```
